On Misuses of the Kolmogorov--Smirnov Test for One-Sample Goodness-of-Fit

### 2023-03-23 manuscript/manuscript.tex
one example reference of misuse
edits on description of working arma

The standard KS test does not apply to stationary yet dependent data either. The
distribution of the KS statistic would have a higher variance for positively
dependent data than that derived when the data are independent because of a
smaller effective sample size. For example, for testing normality,
\citet{durilleul1992lack} demonstrates that a naive application of
the KS statistic is too liberal for medium-to-high positive serial dependence,
and that for negative dependence, the behavior is asymmetrical. For remedies,
\citet{weiss1978modification} provides a procedure that is applicable 
specifically for data
modeled by the second-order auto-regressive (AR) process where the AR parameters
are known. \citet{lanzante2021testing} tests various strategies for dealing with
temporal
dependence and concludes that a test based on Monte-Carlo simulations performed
the best. When additionally the hypothesized distribution contains unknown
parameters, the standard KS test becomes even further inapplicable.

The contribution of this paper is a demonstration of misuses of the one-sample
KS test in three scenarios and their remedies in practice. The scenarios are
where:
1) the hypothesized distribution has unspecified parameters;
2) the data are serially dependent; and
3) a combination of the first two scenarios.
In each scenario, the misuse is performed and the impacts are shown. Then, a
remedy is detailed and performed alongside the misuse to show its positive
effects. In order to set up the demonstrations, simulated data are used
throughout. The remedies are also performed on various families of
distributions. An R implementation of the proposed methods will be available
after the paper is published.

The rest of the paper is organized as follows. Section~\ref{sec:fitted}
investigates the scenario where the hypothesized distribution has unspecified
parameters. Both parametric and nonparametric bootstrap are available to fix the
issue. Section~\ref{sec:dependence} investigates the scenario where the data of
the empirical distribution are serially dependent. A bootstrap procedure
employing copulas to account for dependence is proposed as a working solution.
Section~\ref{sec:fittedwithdependence}
explores the case where a combination of the first two scenarios occurs. The
copula procedure can be adjusted for the use of fitted parameters as a remedy.
Section~\ref{sec:conclusion} concludes with a discussion.

@@

Nonparametric bootstrap can also be used to approximate the null distribution
of the test statistic with the bias correction from \citet{babu2004goodness}:
\begin{enumerate}
\item
  Draw a random sample $X_1^*,...,X_n^*$ from the empirical distribution $F_n$
  with replacement
\item
  Fit $F_\theta$ to $X_1^*,...,X_n^*$ and obtain estimate 
	$\hat\theta_n^*$ of $\theta$.
\item
  Obtain the empirical distribution function $F_n^*$ of
  $X_1^*, \ldots,  X_n^*$. \eds{verify edit}
\item
  Calculate bootstrap KS statistic
  \[
    D_n^* = \sup_x | \sqrt{n}\left(F_n^* (x)- F_{\hat\theta_n^*}(x)\right) - B_n(x) |.
  \]
  where $B_{n}(x) = \sqrt{n}(F_{n}(x) - F_{\hat\theta_n}(x))$ is the known
  bias term \citep{babu2004goodness}.
\item
  Repeat the previous steps a large number $B$ times and use the empirical
  distribution of $D_n^*$ to approximate the null distribution of the observed
  statistic. 
\end{enumerate}
The p-value of $D_n$ is, again, approximated by the proportion of the $D_n^*$
statistics that are greater than or equal to~$D_n$. Note that the step~1 and
step~4 are different from the parametric bootstrap version.

@@

We propose a semiparametric bootstrap procedure that assumes a working serial
dependence structure which does not need to be correctly specified. The working
serial dependence is introduced through copulas with the hope to approximate the
true serial dependence as allowed by the working copulas. A copula is a
multivariate distribution with standard uniform marginal distributions, which
completely characterizes the dependence structure of a multivariate
distribution \citep{hofert2018elements}. It separates the dependence structure
of a multivariate distribution from its marginal distributions. The normal copula
is the copula determined by the multivariate normal distribution. We propose to
use a normal working copula to approximate the serial
dependence of the observations.

The parameters of the working normal copula are
estimated from fitting an auto-regressive moving average (ARMA) model to the
observed data transformed to the standard normal scale. In particular, let
$Z_i = \Phi^{-1}\{ F(X_i)\}$, $i = 1, \ldots, n$, where $\Phi$ is the distribution
function of the standard normal. Then we fit an ARMA$(p, q)$ model to
$Z_1, \ldots, Z_n$ with AR order $p$ and MA order $q$ selected by the Akaike
Information Criterion (AIC). This can be done with, for example, function
\texttt{auto.arima()} from  R package \texttt{forecast}
\citep{hyndman2008automatic}. Since it is known that the mean of $Z_i$'s is
zero, it is necessary to set the intercept of the ARMA model as zero in the
fitting process. This restriction turned out to be critical in our
investigation; having the intercept estimated does not lead to desired p-value
distributions in the following bootstrap process.
The selected ARMA$(p, q)$ model with fitted
parameters will be used as the working model to generate bootstrap samples with
serial dependence mimicking that in the observed data. The unconditional
variance $\sigma^2$ of the ARMA model with unit innovation variance can be
obtained with function
\texttt{tacvfARMA()} from R package \texttt{ltsa} \citep{mcleod2007algorithms}.

The semiparametric bootstrap procedure is as follows.
\begin{enumerate}
\item
  Generate $Z_1^*, \ldots, Z_n^*$ from the fitted ARMA$(p, q)$ process with
  innovation \eds{? inverse?} variance $1 / \sigma^2$ such that the $Z_i$'s are
	marginally \jy{yes, it means the innovation series of an ARMA process.}
  standard normal variables.
\item
  Form a bootstrap sample $X_i^* = F^{-1} [\Phi(Z_i^*)]$,  $i = 1, \ldots, n$.
\item
  Obtain the empirical distribution function $F_n^*$ of the bootstrap sample
  $X_1^*, \ldots, X_n^*$.
\item
  Calculate bootstrap KS statistic
  \[
    D_n^* = \sup_x \lvert F_n^* (x)- F(x) \rvert.
  \]
\item
  Repeat the previous steps a large number $B$ times and use the empirical
  distribution of the $B$ test statistics to approximate
  the null distribution of the observed statistic.
\end{enumerate}
The p-value of $D_n$ is, again, approximated by the proportion of the $D_n^*$
statistics that are greater than or equal to~$D_n$.

@@

Specifically, let $\hat\theta_n$ be the marginally fitted parameters, and
$Z_i = \Phi^{-1} \{F_{\hat\theta_n}(X_i)\}$, $i = 1, \ldots, n$ be a
transformation of the observed data onto the standard normal scale using the
fitted parameters $\hat\theta_n$. We then fit an ARMA model with orders
selected by the AIC using function \texttt{auto.arima()} from R package
\texttt{forecase}. This ARMA model will be used as the working model to
introduce serial dependence in the bootstrap samples. Similarly, we obtain the
unconditional variance $\sigma^2$ of the ARMA process with unit innovation
variance. Our semiparametric bootstrap procedure is as follows

\begin{enumerate}
\item
  Generate $Z_1^*, \ldots, Z_n^*$ from the working ARMA$(p, q)$ model with
  innovation variance $1 / \sigma^2$ such that the $Z_i$'s are marginally
  standard normal variables.
\item
  Form a bootstrap sample
  $X_i^* = F_{\hat\theta_n}^{-1} [\Phi(Z_i^*)]$, $i = 1, \ldots, n$.
\item
  Fir $F_\theta$ to $X_1^*, \ldots, X_n^*$ to obtain estimate $\hat\theta_n^*$
  of $\theta$.
\item
  Obtain the empirical distribution function $F_n^*$ of the bootstrap sample
  $X_1^*, \ldots, X_n^*$.
\item
  Calculate bootstrap KS statistic
  \[
    D_n^* = \sup_x \lvert F_n^* (x) - F_{\hat\theta_n}^* (x) \rvert.
  \]
\item
  Repeat the previous steps a large number $B$ times and use the empirical
  distribution of the $B$ test statistics to approximate
  the null distribution of the observed statistic.
\end{enumerate}
The p-value of $D_n$ is approximated by the proportion of the $D_n^*$
statistics that are greater than or equal to~$D_n$. Compared to the algorithm in
the last section, here we use $F_{\hat\theta_n}$ and $F_{\hat\theta_n}^*$ in
place of the known~$F$.

>>>>

The standard KS test does not apply to stationary yet dependent data either. The
distribution of the KS statistic would have a higher variance for positively
dependent data than that derived when the data are independent because of a
smaller effective sample size. For example, for testing normality,
\citet{durilleul1992lack} demonstrates that a naive application of
the KS statistic is too liberal for medium-to-high positive serial dependence,
and that for negative dependence, the behavior is asymmetrical. For remedies,
\citet{weiss1978modification} provides a procedure that is applicable 
specifically for data
modeled by the second-order auto-regressive (AR) process where the AR parameters
are known. \citet{lanzante2021testing} tests various strategies for dealing with
temporal
dependence and concludes that a test based on Monte-Carlo simulations performed
the best. When additionally the hypothesized distribution contains unknown
parameters, the standard KS test becomes even further
inapplicable. Practitioners, however, may have used the test without knowing its
limitations \citep[e.g.,][]{tuncer2019online}.

The contribution of this paper is a demonstration of misuses of the one-sample
KS test in three scenarios and their remedies in practice. The scenarios are
where:
1) the hypothesized distribution has unspecified parameters;
2) the data are serially dependent; and
3) a combination of the first two scenarios.
In each scenario, the misuse is performed and the impacts are shown. Then, a
remedy is detailed and performed alongside the misuse to show its positive
effects. Specifically, unspecified parameters are handled by parametric
bootstrap; serial dependence is handled by introducing a working autoregressive
moving average (ARMA) model that preserves the serial dependence for a wide
range of dependence structures.
In order to set up the demonstrations, simulated data are used
throughout. The remedies are also performed on various families of
distributions. An R implementation of the proposed methods will be available
after the paper is published.

The rest of the paper is organized as follows. Section~\ref{sec:fitted}
investigates the scenario where the hypothesized distribution has unspecified
parameters. Both parametric and nonparametric bootstrap are available to fix the
issue. Section~\ref{sec:dependence} investigates the scenario where the data of
the empirical distribution are serially dependent. An bootstrap procedure
employing a working ARMA model to account for dependence is proposed as a
working solution. Section~\ref{sec:fittedwithdependence}
explores the case where a combination of the first two scenarios occurs.
Section~\ref{sec:conclusion} concludes with a discussion.

@@

Nonparametric bootstrap can also be used to approximate the null distribution
of the test statistic with the bias correction from \citet{babu2004goodness}:
\begin{enumerate}
\item
  Draw a random sample $X_1^*,...,X_n^*$ from the empirical distribution $F_n$
  with replacement
\item
  Fit $F_\theta$ to $X_1^*,...,X_n^*$ and obtain estimate 
	$\hat\theta_n^*$ of $\theta$.
\item
  Obtain the empirical distribution function $F_n^*$ of
  $X_1^*, \ldots,  X_n^*$. \eds{verify edit} \jy{Good.}
\item
  Calculate bootstrap KS statistic
  \[
    D_n^* = \sup_x | \sqrt{n}\left(F_n^* (x)- F_{\hat\theta_n^*}(x)\right) - B_n(x) |.
  \]
  where $B_{n}(x) = \sqrt{n}(F_{n}(x) - F_{\hat\theta_n}(x))$ is the known
  bias term \citep{babu2004goodness}.
\item
  Repeat the previous steps a large number $B$ times and use the empirical
  distribution of $D_n^*$ to approximate the null distribution of the observed
  statistic. 
\end{enumerate}
The p-value of $D_n$ is, again, approximated by the proportion of the $D_n^*$
statistics that are greater than or equal to~$D_n$. Note that the step~1 and
step~4 are different from the parametric bootstrap version.

@@

We propose a semiparametric bootstrap procedure that assumes a working serial
dependence structure which does not need to be correctly specified. The working
serial dependence is introduced through a working ARMA process with the hope to
approximate the true serial dependence as allowed by the working
model. Essentially, this working model covers the most commonly seen dependence
structure characterized by a normal copula \citep{hofert2018elements}, which
separates the dependence structure
of a multivariate distribution from its marginal distributions. 
The parameters of the working normal copula are
estimated from fitting an ARMA model to the
observed data transformed to the standard normal scale.

In particular, let
$Z_i = \Phi^{-1}\{ F(X_i)\}$, $i = 1, \ldots, n$, where $\Phi$ is the distribution
function of the standard normal. Then we fit an ARMA$(p, q)$ model to
$Z_1, \ldots, Z_n$ with AR order $p$ and MA order $q$ selected by the Akaike
Information Criterion (AIC). This can be done with, for example, function
\texttt{auto.arima()} from  R package \texttt{forecast}
\citep{hyndman2008automatic}. Since it is known that the mean of $Z_i$'s is
zero, it is necessary to set the intercept of the ARMA model as zero in the
fitting process. This restriction turned out to be critical in our
investigation; having the intercept estimated does not lead to desired p-value
distributions in the following bootstrap process.
The selected ARMA$(p, q)$ model with fitted
parameters will be used as the working model to generate bootstrap samples with
serial dependence mimicking that in the observed data. The unconditional
variance $\sigma^2$ of the ARMA model with unit innovation variance can be
obtained with function
\texttt{tacvfARMA()} from R package \texttt{ltsa} \citep{mcleod2007algorithms}.

The semiparametric bootstrap procedure is as follows.
\begin{enumerate}
\item
  Generate $Z_1^*, \ldots, Z_n^*$ from the fitted ARMA$(p, q)$ process with
  innovation \eds{? inverse?} variance $1 / \sigma^2$ such that the $Z_i$'s are
	marginally \jy{yes, it means the innovation series of an ARMA process.}
  standard normal variables.
\item
  Form a bootstrap sample $X_i^* = F^{-1} [\Phi(Z_i^*)]$,  $i = 1, \ldots, n$,
  where $F^{-1}$ is the quantile function of $F$.
\item
  Obtain the empirical distribution function $F_n^*$ of the bootstrap sample
  $X_1^*, \ldots, X_n^*$.
\item
  Calculate bootstrap KS statistic
  \[
    D_n^* = \sup_x \lvert F_n^* (x)- F(x) \rvert.
  \]
\item
  Repeat the previous steps a large number $B$ times and use the empirical
  distribution of the $B$ test statistics to approximate
  the null distribution of the observed statistic.
\end{enumerate}
The p-value of $D_n$ is, again, approximated by the proportion of the $D_n^*$
statistics that are greater than or equal to~$D_n$.

@@

Specifically, let $\hat\theta_n$ be the marginally fitted parameters, and
$Z_i = \Phi^{-1} \{F_{\hat\theta_n}(X_i)\}$, $i = 1, \ldots, n$ be a
transformation of the observed data onto the standard normal scale using the
fitted parameters $\hat\theta_n$. We then fit an ARMA model with orders
selected by the AIC using function \texttt{auto.arima()} from R package
\texttt{forecast}. This ARMA model will be used as the working model to
introduce serial dependence in the bootstrap samples. Again, we obtain the
unconditional variance $\sigma^2$ of the ARMA process with unit innovation
variance. Our semiparametric bootstrap procedure is as follows.

\begin{enumerate}
\item
  Generate $Z_1^*, \ldots, Z_n^*$ from the working ARMA$(p, q)$ model with
  innovation variance $1 / \sigma^2$ such that the $Z_i$'s are marginally
  standard normal variables.
\item
  Form a bootstrap sample
  $X_i^* = F_{\hat\theta_n}^{-1} [\Phi(Z_i^*)]$, $i = 1, \ldots, n$, where
  $F_{\hat\theta_n}^{-1}$ is the quantile function of $F_{\hat\theta_n}$.
\item
  Fir $F_\theta$ to $X_1^*, \ldots, X_n^*$ to obtain estimate $\hat\theta_n^*$
  of $\theta$.
\item
  Obtain the empirical distribution function $F_n^*$ of the bootstrap sample
  $X_1^*, \ldots, X_n^*$.
\item
  Calculate bootstrap KS statistic
  \[
    D_n^* = \sup_x \lvert F_n^* (x) - F_{\hat\theta_n}^* (x) \rvert.
  \]
\item
  Repeat the previous steps a large number $B$ times and use the empirical
  distribution of the $B$ test statistics to approximate
  the null distribution of the observed statistic.
\end{enumerate}
The p-value of $D_n$ is approximated by the proportion of the $D_n^*$
statistics that are greater than or equal to~$D_n$. Compared to the algorithm in
the last section, here we use $F_{\hat\theta_n}$ and $F_{\hat\theta_n}^*$ in
place of the known~$F$.
